---
title: "Project 1"
author: "Eya MAHDHAOUI ,Manon CASTANET,Aline RIVERA, Pedro José TRILLOS"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = "allow")
```

### Introduction:  
In this project we aim to calculate the probability that the price of coinbase cryptocurrency is lower by at least 9% on at least one of the next 5 trading days. 

#### 1- Import data: 
```{r , echo= TRUE }
data = read.csv("bitcoinity_data.csv")
View(data)
```  

#### 2- Clean the data:  
We removed the missing values to be able to calculate the mean value.  

```{r , echo= TRUE }
coinbase_price= (data$coinbase)
coinbase_price= na.omit(coinbase_price)
```  
#### 3- Calculation of the mean and standard deviation:  
In order to caclculate the probability that the coinbase cryptocurrency will be lower by at least 9% on at least one of the next 5 trading days, we need to calculate first the mean and the standard deviation of the sample data.  

```{r , echo = TRUE}
m= mean(coinbase_price)
std = sd(coinbase_price)
p =coinbase_price[2504]
```  

#### 4- Calcutation of the probability:  
We created a vector called "below" which contains 100000 zeros in order to put 
```{r , echo = TRUE}
below = rep(0, 100000)
set.seed(2000)
for (i in 1:100000) {
  r = rnorm(5,m,std) 
  Price = cumsum(r) 
  minP = min(Price) 
  below[i] = as.numeric(minP - p < 0,09)
}
```  

```{r , echo = TRUE}
mean(below)
```  

#### Conclusion:  
The probability that the price of coinbase cryptocurrency is lower by at least 9% on at least one of the next 5 trading days is equal to 68%.In other words, an investor in coinbase cryptocurrency will be likely to lose at least 9% if he decided to sell in the next 5 trading days.    




